---
output:
  md_document:
    variant: markdown_github
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  cache = TRUE
)
```
<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/tgstat)](https://CRAN.R-project.org/package=tgstat)
<!-- badges: end -->

# tgstat

The goal of `tgstat` is to provide fast and efficient implementation of certain R functions
such as 'cor' and 'dist', along with specific statistical tools.

Various approaches are used to boost the performance, including multi-processing and use of
optimized functions provided by the Basic Linear Algebra Subprograms (BLAS) library.

## Installation

Install from CRAN: 

```{r, eval = FALSE}
install.packages("tgstat")
```

For the development version:

```{r, eval = FALSE}
remotes::install_github("tanaylab/tgstat")
```

## Examples

```{r, eval=TRUE}
set.seed(seed=60427)
rows = 3000
cols = 3000
vals <- sample(1:(rows*cols/2), rows*cols, replace=T)
m <- matrix(vals, nrow=rows, ncol=cols)
m_with_NAs <- m
m_with_NAs[sample(1:(rows*cols), rows*cols / 10)] <- NA
dim(m)
```

### Fast computation of correlation matrices

Pearson correlation without BLAS, no NAs:
 
```{r, eval = TRUE}
options(tgs_use.blas=F)
system.time(tgs_cor(m))
```

Same with BLAS:
 
```{r, eval = TRUE}
# tgs_cor, with BLAS, no NAs, pearson
options(tgs_use.blas=T)
system.time(tgs_cor(m))
```

Base R version:

```{r, eval = TRUE}
system.time(cor(m))
```

Pearson correlation without BLAS, with NAs:
 
```{r, eval = TRUE}
options(tgs_use.blas=F)
system.time(tgs_cor(m_with_NAs, pairwise.complete.obs=T))
```

Same with BLAS:
 
```{r, eval = TRUE}
options(tgs_use.blas=T)
system.time(tgs_cor(m_with_NAs, pairwise.complete.obs=T))
```

Base R version:

```{r, eval = TRUE}
system.time(cor(m_with_NAs, use="pairwise.complete.obs"))
```

### Fast computation of distance matrices

Distance without BLAS, no NAs:

```{r, eval = TRUE}
options(tgs_use.blas=F)
system.time(tgs_dist(m))
```

Same with BLAS:

```{r, eval = TRUE}
options(tgs_use.blas=T)
system.time(tgs_dist(m))
```

Base R:

```{r, eval = TRUE}
system.time(dist(m, method="euclidean"))
```

## Notes regarding the usage of `BLAS`

`tgstat` runs best when R is linked with an optimized BLAS implementation.

Many optimized BLAS implementations are available, both proprietary (e.g. Intel's MKL,
Apple's vecLib) and opensource (e.g. OpenBLAS, ATLAS). Unfortunately, R often uses by default
the reference BLAS implementation, which is known to have poor performance.

Having `tgstat` rely on the reference BLAS will result in very poor performance and is strongly
discouraged. If your R implementation uses an optimized BLAS, set `options(tgs_use.blas=TRUE)`
to allow `tgstat` to make BLAS calls. Otherwise, set `options(tgs_use.blas=FALSE)` (default)
which instructs `tgstat` to avoid BLAS and instead rely only on its own optimization methods.
If in doubt, it is possible to run one of `tgstat` CPU intensive functions (e.g. `tgs_cor`) and
compare its run time under both `options(tgs_use.blas=FALSE)`.

Exact instructions for linking R with an optimized BLAS library are system dependent and are out
of scope of this document.